---
title: "ADHD management"
format:
  html:
    embed-resources: true
    toc: true
params:
  data_path: ""    
---


# ADHD Analysis

```{r setup_height}

figure_height <- 10



```

```{r init}
#| results: 'asis'
#| fig-width: 8
#| fig-height: !expr figure_height

# ADHD Analysis 2025-04-23
library(broom)
library(dplyr)
library(glue)
library(htmltools)
library(knitr)
library(kableExtra)
library(metafor)



# Load using the absolute path passed from the main script
if (params$data_path != "") {
  load(params$data_path)
}


# Remove empty intervention
d <- d %>% dplyr::filter(intervention != "")


# find the continuous variable intervention names
int_continuous_vars <- 
  d %>% 
  dplyr::select(ends_with("mean_int_cont")) %>% 
  names() %>%
  gsub("_.+", "", .)

ctrl_continuous_vars <- 
  d %>% 
  dplyr::select(ends_with("mean_ctrl_cont")) %>% 
  names() %>%
  gsub("_.+", "", .)



list_variable_types <-
  list(
    c(var = "functional", type = "continuous"),
    c(var = "functional", type = "categorical"),
    c(var = "cognitive", type = "continuous"),
    c(var = "cognitive", type = "categorical"),
    c(var = "symptom", type = "continuous"),
    c(var = "symptom", type = "categorical"),
    c(var = "substance_use", type = "continuous"),
    c(var = "substance_use", type = "categorical"),
    c(var = "relationship", type = "continuous"),
    c(var = "relationship", type = "categorical"),
    c(var = "BP", type = "continuous"),
    c(var = "BP", type = "categorical"),
    c(var = "sleep_disturbance", type = "continuous"),
    c(var = "sleep_disturbance", type = "categorical"),
    c(var = "AEs", type = "continuous"),
    c(var = "AEs", type = "categorical"),
    c(var = "mortality", type = "continuous"),
    c(var = "mortality", type = "categorical"),    
    c(var = "behavior", type = "continuous"),
    c(var = "behavior", type = "categorical"),
    c(var = "appetite", type = "continuous"),
    c(var = "appetite", type = "categorical"),
    c(var = "mood", type = "continuous"),
    c(var = "mood", type = "categorical")
  )


int_continuous_vars <- 
  ifelse(int_continuous_vars == "substance", 
         "substance_use", 
         int_continuous_vars
  ) 
int_continuous_vars <- 
  ifelse(int_continuous_vars == "sleep", 
         "sleep_disturbance", 
         int_continuous_vars
  ) 

if (length(int_continuous_vars) != length(ctrl_continuous_vars)) {
  stop("Problem with continuous variables")
}



# find the categorical variable outcome names
int_categ_vars <- 
  d %>% 
  dplyr::select(ends_with("int_categ")) %>% 
  names() %>%
  gsub("_.+", "", .) %>% 
  unique()

ctrl_categ_vars <- 
  d %>% 
  dplyr::select(ends_with("ctrl_categ")) %>% 
  names() %>%
  gsub("_.+", "", .) %>% 
  unique()

int_categ_vars <- 
  ifelse(int_categ_vars == "substance", "substance_use", int_categ_vars) 
int_categ_vars <- 
  ifelse(int_categ_vars == "sleep", "sleep_disturbance", int_categ_vars) 

if (length(int_categ_vars) != length(ctrl_categ_vars)) {
  stop("Problem with continuous variables")
}

# need to rename substance to substance_use


# Check the direction for each outcome that we want for the effects and 
# forest plots.
# Variable says what we want on the right hand side of the x-axis.
# e.g. blood pressure lower <-----> higher,
# cognitive higher <---------> lower
d <-  d %>% 
  dplyr::mutate(
    # continuous directions 
    AEs_positive_direction = "higher",
    appetite_positive_direction =  "higher",   ######### Check
    behavior_positive_direction = "lower",
    BP_positive_direction = "higher",
    cognitive_positive_direction = "lower",
    functional_positive_direction = "higher",
    mood_positive_direction = "higher",   ######### Check
    mortality_positive_direction = "lower",
    relationship_positive_direction = "higher",
    sleep_disturbance_positive_direction = "higher",
    substance_use_positive_direction = "lower",
    symptom_positive_direction = "higher",
    # categorical
    AEs_positive_direction_categ = "higher",
    appetite_positive_direction_categ = "lower",
    behavior_positive_direction_categ = "lower",
    BP_positive_direction_categ = "lower",
    cognitive_positive_direction_categ = "higher", 
    functional_positive_direction_categ = "lower",
    mood_positive_direction_categ = "lower",     # check
    mortality_positive_direction_categ = "lower",
    relationship_positive_direction_categ = "lower",
    sleep_disturbance_positive_direction_categ = "higher",
    substance_use_positive_direction_categ = "higher",
    symptom_positive_direction_categ = "higher"
  )




```

Read the functions to do MA:

```{r read_functions}

source(glue::glue(git_raw_root, "/helper_functions/AmphMphMetaReg.R"))
source(glue::glue(git_raw_root, "/helper_functions/CatMA.R"))
source(glue::glue(git_raw_root, "/helper_functions/CombinedMetaregs.R"))
source(glue::glue(git_raw_root, "/helper_functions/contMA.R"))
source(glue::glue(git_raw_root, "/helper_functions/MPH_Dose_metareg.R"))
source(glue::glue(git_raw_root, "/helper_functions/StimMetaReg.R"))
```



```{r}
#| results: 'asis'
#| fig-width: 8
#| fig-height: !expr figure_height

d <- d %>%
  dplyr::mutate(
    symptom_SD_ctrl_cont = ifelse(symptom_SD_ctrl_cont  < 0, NA, symptom_SD_ctrl_cont),
    symptom_SD_int_cont = ifelse(symptom_SD_int_cont  < 0, NA, symptom_SD_int_cont),
  )


```

# All interventions 

```{r all_interventions}
#| results: 'asis'
#| #| fig-width: 8
#| fig-height: !expr figure_height


for (var_num in 1:length(list_variable_types)) {
  current_intervention <- ""
  outcome_name <- list_variable_types[[var_num]]["var"]
  outcome_type <- list_variable_types[[var_num]]["type"]
  
  cat(
    glue::glue("<h2> Intervention: All Interventions <br> 
                <h3>Outcome: {outcome_name}:{outcome_type}</h3><p>")
  )
  
  if (outcome_type == "continuous") {
    res <- ContMA(
      data = d,
      label = "ID",
      outcome = outcome_name,
      comparator_vs_int = comparator_vs_int
    ) 
  }
  if (outcome_type == "categorical") {
    res <- CatMA(
      data = d,
      label = "ID",
      outcome = outcome_name,
      comparator_vs_int = comparator_vs_int
    )
  }
}

```


# Break down by interventions

```{r by_intervention}
#| results: 'asis'
#| fig-width: 8
#| fig-height: !expr figure_height

if (subgroup == "") {
  for (current_intervention in sort(unique(d$intervention))) {
    cat(glue::glue("<h1> Intervention: {current_intervention}</h1> "))
    cat("xxxxxxx")    
    medication_analysis <- current_intervention == "Medication"
    # If current_intervention is medication, we're going to also use 
    # comparator as an intervention group, as long as the intervention is not
    # also medication
    # if doing medication analysis, we also want medication in comparator
    if (!medication_analysis) {
      d_short <- d %>%
        dplyr::filter(
          intervention == current_intervention | 
            comparator == current_intervention, 
        )
    } else {
      d_short <- d %>%
        dplyr::filter(
          intervention == current_intervention | 
            comparator == current_intervention, 
        )
    }
    
    for (var_num in 1:length(list_variable_types)) {
      
      outcome_name <- list_variable_types[[var_num]]["var"]
      outcome_type <- list_variable_types[[var_num]]["type"]
      
      cat(
        glue::glue(
          "<h3>Outcome: {outcome_name}:{outcome_type} 
          Intervention: {current_intervention}</h3><p>")
      )
      
      if (outcome_type == "continuous") {
        res <- ContMA(
          data = d_short,
          label = "ID",
          outcome = outcome_name,
          comparator_vs_int = comparator_vs_int
        ) 
      }
      if (outcome_type == "categorical") {
        print(glue::glue("Outcome name: {outcome_name}"))
#        if (outcome_name == "symptom" & current_intervention == "cognitive") {
#          stop("Got to cog cat")
#        }
        res <- CatMA(
          data = d_short,
          label = "ID",
          outcome = outcome_name,
          comparator_vs_int = comparator_vs_int,
          medication_analysis = medication_analysis
        )

      }
    }
  }
}

```

# Subintervention  Analysis
Perform meta-analyses separately for MPH, AMPH, and NRI's on ADHD symptoms, cognitive performance, and adverse events (p.28)? Reviewers and readers will absolutely expect and want it.

int_methylphenidate
int_amphetamine
int_NRI

•	Perform meta-analyses separately for MBCT, DBT, and CBT on ADHD symptoms (p.33)?
- int_mindfulness
- int_DBT
- int_CBT_only


Continuous Outcomes:
- cognitive  symptom AEs
Categorical outcomes:
- cognitive symptom AEs


```{r mph_amph_nri}
#| results: 'asis'
#| fig-width: 8
#| fig-height: !expr figure_height

#Only doing specific interventions, and only specific outcomes.



sub_interventions <- 
  c("longterm", 
    "int_stimulant", "int_nonstimulant", 
    "int_methylphenidate", "int_amphetamine", 
    "int_NRI", "int_alpha_agonist", "int_atomoxetine",
    "int_bupropion", 
    "int_mindfulness", "int_DBT_only", "int_CBT_only", "int_MBCT_only"
    )


rm(d_short)

for (current_intervention in sort(unique(sub_interventions))) {

  medication_analysis <- FALSE
  if (current_intervention %in% 
      c("int_methylphenidate", "int_amphetamine", "int_NRI",
        "int_stimulant",
        "int_nonstimulant")) {
    medication_analysis <- TRUE
  }
  
  d_short <- d[d[current_intervention] == current_intervention, ]
  
  medication_analysis <- FALSE
  if (current_intervention %in% 
      c("int_methylphenidate", "int_amphetamine", "int_NRI",
        "int_stimulant",
        "int_nonstimulant")) {
    medication_analysis <- TRUE
  }
  
  
  cat(glue::glue("<h2> MPH, AMPH, and NRI's Section: 
                 Intervention: {current_intervention}</h2> "))
  for (var_num in 1:length(list_variable_types)) {
    
    outcome_name <- list_variable_types[[var_num]]["var"]
    outcome_type <- list_variable_types[[var_num]]["type"]
    
    cat(
      glue::glue("<h2> Intervention: {current_intervention} <br> 
                <h3>Outcome: {outcome_name}:{outcome_type}</h3><p>")
    )
    
    if (outcome_type == "continuous") {
      res <- ContMA(
        data = d_short,
        label = "ID",
        outcome = outcome_name,
        comparator_vs_int = comparator_vs_int
      ) 
    }
    if (outcome_type == "categorical") {
      res <- CatMA(
        data = d_short,
        label = "ID",
        outcome = outcome_name,
        comparator_vs_int = comparator_vs_int
      )
    }
  }
}  

rm(d_short)

```


# atomoxetine vs methylphenidate


```{r atom_vs_meth}
#| results: 'asis'
#| fig-width: 8
#| fig-height: !expr figure_height


d_short <- d %>%
  dplyr::filter(
    (int_methylphenidate == "int_methylphenidate" & comp_atomoxetine == "comp_atomoxetine") |
      (comp_methylphenidate == "comp_methylphenidate" & int_atomoxetine == "int_atomoxetine")
  )
nrow(d_short)


d_short %>% dplyr::group_by(
  comp_methylphenidate, int_atomoxetine,
  comp_atomoxetine,  int_methylphenidate
) %>% dplyr::summarise(n_studies = dplyr::n()) %>%
  as.data.frame() %>% 
  knitr::kable()

current_intervention <- "Atomoxetine vs Methylphenidate"

cat(glue::glue("There are three studies for methylphenidate, and in each study 
int is atomoxetine, and comp is always methylphenidate. That
means we don't need to do difficult swaps, but we need to 
remember that comp is methylphenidate") )




#Only doing specific interventions, and only specific outcomes.

medication_analysis <- FALSE


cat(glue::glue("<h2> MPH, AMPH, and NRI's Section: 
                 Intervention: {current_intervention}</h2> "))
for (var_num in 1:length(list_variable_types)) {
  
  outcome_name <- list_variable_types[[var_num]]["var"]
  outcome_type <- list_variable_types[[var_num]]["type"]
  
  cat(
    glue::glue("<h2> Intervention: {current_intervention} <br> 
                <h3>Outcome: {outcome_name}:{outcome_type}</h3><p>")
  )
  
  if (outcome_type == "continuous") {
    res <- ContMA(
      data = d_short,
      label = "ID",
      outcome = outcome_name,
      comparator_vs_int = TRUE
    ) 
  }
  if (outcome_type == "categorical") {
    res <- CatMA(
      data = d_short,
      label = "ID",
      outcome = outcome_name,
      comparator_vs_int = TRUE
    )
  }
}

```





# amphetamine  vs methylphenidate

```{r amph_vs_meth}
#| results: 'asis'
#| fig-width: 8
#| fig-height: !expr figure_height


d_short <- d %>%
  dplyr::filter(
    (int_methylphenidate == "int_methylphenidate" & comp_atomoxetine == "comp_amphetamine") |
      (comp_methylphenidate == "comp_methylphenidate" & int_atomoxetine == "int_atomoxetine")
  )
nrow(d_short)


d_short %>% dplyr::group_by(
  comp_methylphenidate, int_atomoxetine,
  comp_atomoxetine,  int_methylphenidate
) %>% dplyr::summarise(n_studies = dplyr::n()) %>%
  as.data.frame() %>% 
  knitr::kable()

current_intervention <- "Atomoxetine vs Methylphenidate"

cat(glue::glue("There are three studies for methylphenidate, and in each study 
int is atomoxetine, and comp is always methylphenidate. That
means we don't need to do difficult swaps, but we need to 
remember that comp is methylphenidate") )

#Only doing specific interventions, and only specific outcomes.
medication_analysis <- FALSE

cat(glue::glue("<h2> MPH, AMPH, and NRI's Section: 
                 Intervention: {current_intervention}</h2> "))
for (var_num in 1:length(list_variable_types)) {
  
  outcome_name <- list_variable_types[[var_num]]["var"]
  outcome_type <- list_variable_types[[var_num]]["type"]
  
  cat(
    glue::glue("<h2> Intervention: {current_intervention} <br> 
                <h3>Outcome: {outcome_name}:{outcome_type}</h3><p>")
  )
  
  if (outcome_type == "continuous") {
    res <- ContMA(
      data = d_short,
      label = "ID",
      outcome = outcome_name,
      comparator_vs_int = TRUE
    ) 
  }
  if (outcome_type == "categorical") {
    res <- CatMA(
      data = d_short,
      label = "ID",
      outcome = outcome_name,
      comparator_vs_int = TRUE
    )
  }
}

```





